Reproducible Research 1
=========

```{r}
library(plyr)
```
Load the raw data..
```{r}
actData <- read.csv('e:/ReproduceResearch/activity.csv',header=T,sep=',')
```
Collapse the 5 minute interval data into daily totals.
```{r}
dailyData <- ddply(actData, c('date'),summarize, dailySteps=sum(steps,na.rm=T))
```
## What is the mean number of steps taken per day?
```{r}
hist(dailyData$dailySteps,main='Daily Steps per Day',xlab='Total Steps per Day',
      ylim=range(0:30))
summary(dailyData$dailySteps)
```

## What is the average daily activity pattern?
Calculate the average number of steps for each 5 minute intervarl over the 61 days of data.
```{r}
minuteData <- ddply(actData,c('interval'),summarize,meanSteps=mean(steps,na.rm=T))
```
Plot the daily activity pattern and show which period was the most active.
```{r}
plot(minuteData$interval,minuteData$meanSteps,type='l',xlab='5 minute time period',
      ylab='Mean # of Steps per Period', main="Daily Acitivy Pattern")
```
Find the most active period during the day.
```{r}
minuteData[which.max(minuteData$meanSteps),]
```
## How much data is missing?
```{r}
sum(is.na(actData$steps))
```
## Replace the missing interval data with the mean number of steps for that interval
```{r}
ad2 <- actData
for (k in 1:length(ad2$steps)) {
      if (is.na(ad2$steps[k])) {
            iTime <- (ad2$interval[k] + 5) / 5
            ad2$steps[k]  <- minuteData$meanSteps[iTime]
      }
}
```
Find the new daily total number of steps taken.
```{r}
dd2 <- ddply(ad2, c('date'),summarize, dailySteps=sum(steps,na.rm=T))
```
Compare the two steps of data.
```{r}
par(mfcol=c(2,1))
hist(dailyData$dailySteps,main='Daily Steps per Day with Missing',xlab='Total Steps per Day',
      ylim=range(0:40),ylab='Frequency')
hist(dd2$dailySteps,main='Daily Steps per Day with Missing Replaced',xlab='Total Steps per Day',
      ylim=range(0:40))
```

## Are there differences in activity patterns between weekdays and weekends?
Add day name to the orginal data set.
```{r}
ad2$dayName <- weekdays(as.Date(ad2$date))
```
Determine if day falls on the weekend or not.
```{r}
for (k in 1:length(ad2$date)) {
      if (ad2$dayName[k] == 'Sunday' | ad2$dayName[k] == 'Saturday') {
            ad2$dayType[k] <- "Weekend"
      } 
      else {
            ad2$dayType[k] <- "Workday"
      }
}
```
Split off Monday - Friday.
```{r}
work <- subset(ad2,dayType=="Workday")
workAct <- ddply(work,c('interval'),summarize,meanSteps=mean(steps,na.rm=T))
```
Split off the Weekend.
```{r}
weekend <- subset(ad2,dayType=="Weekend")    
weekendAct <- ddply(weekend,c('interval'),summarize,meanSteps=mean(steps,na.rm=T))
```
Compare the activity patterns between weekdays and weekends.
```{r}
par(mfcol = c(2,1))      

with(workAct,plot(interval,meanSteps,type='l',main='Workday Activity',
                  xlab='Interval (HHmm)',
                  ylab='Mean # of Steps', ylim=range(0:250))
     )
     
with(weekendAct,plot(interval,meanSteps,type='l',main='Weekend Activity',
                     xlab='Interval (HHmm)',
                     ylab='Mean # of Steps', ylim = range(0:250))
     )                       
```